---
title: "IODS Course - Final Assignment"
author: "Tiina Autio"
date: "08/03/2017"
output: html_document
---
###### tiina.autio@helsinki.fi


# Analysis of the library loan data

This data is combination of datasets from Pirkanmaa libraries most loaned 1000 pieces per year from years 2012-2015. To combine all data from all years magazines has been removed from data sets. Original data source (http://www.tampere.fi/tampereen-kaupunki/tietoa-tampereesta/avoin-data.html) and data wrangling file (https://github.com/suprajohde/IODS-final/blob/master/Data/Data%20Wrangling.R) have deeper description of the data.

Let's take a look of the data:
```{r}
library(dplyr)
library_data <- read.csv("~/Documents/IODS-final/Data/library_data.csv", sep=",", header=TRUE, stringsAsFactors = FALSE)
keep_columns <- c("Hyllypaikka", "Tekijä", "Nimeke", "Lainojen.lkm.", "ISBN.ISSN", "Julkaisuvuosi", "Vuosi")
library_data <- select(library_data, one_of(keep_columns))
colnames(library_data)
str(library_data)
glimpse(library_data)
```
Data set has seven variables:  
- shelf location  
- author  
- name
- quantity of the loans  
- ISBN.ISSN id  
- publishing year  
- dataset year  

Dataset consists of 3755 different observations which means that 245 magazines has been removed from original data.

This is the summary of the data:
```{r}
summary(library_data)
```
Summary of the data shows that there are maximum loan quantity has been 3845 and the lowest piece on the list has 425 loans. 